---
title: "Contributing to charlatan"
author: "Scott Chamberlain"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: readable
vignette: >
  %\VignetteIndexEntry{Contributing to charlatan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

`charlatan` is a wee bit complex. This vignette aims to help you contribute 
to the package. 

Let's start with some definitions.

## Definitions

For the purposes of this package:

* **Provider**: a type of data that can be generated in `charlatan`. This may involve
a single file, more than one file, a single R6 class or many R6 classes. 
* **Locale**: a locale for our purposes is a specific spoken language that's 
associated with a specific country. You can have more than one locale for a 
given language (e.g., `en-US`, `en-GB`). Some fakers won't have any locales, 
whereas others can have many.

## Communication

It's a good idea to open an issue if you want to add a new provider or locale to 
an existing provider; it should help make sure there's no duplicated effort and 
we can help make sure you have the knowledge you need.

## Adding a new provider

Providers are generally first created by making an R6 class. Let's start with a 
heavily simplified base R6 class that defines some utility methods. We 
call it `BaseProvider` in `charlatan`, but here we'll call it `MyBaseProvider`
to avoid confusion.

```{r}
library(R6)
MyBaseProvider <- R6::R6Class(
  'MyBaseProvider',
  public = list(
    random_element = function(x) {
      if (length(x) == 0) return('')
      if (inherits(x, "character")) if (!any(nzchar(x))) return('')
      x[sample.int(n = length(x), size = 1)]
    },

    random_int = function(min = 0, max = 9999, size = 1) {
      stopifnot(max >= min)
      num <- max - min + 1
      sample.int(n = num, size = size, replace = TRUE) + (min - 1)
    }
  )
)
```

### Providers without locale support

If you don't need to handle locales it becomes simpler:


```{r}
FooBar <- R6::R6Class(
  'FooBar',
  inherit = BaseProvider,
  public = list(
    integer = function(n = 1, min = 1, max = 1000) {
      super$random_int(min, max, n)
    }
  )
)
```

We can create an instance of the `FooBar` class by calling `$new()` on it.
It only has one method `integer()`, which we can call to get a random 
integer.

```{r}
x <- FooBar$new()
x
x$integer()
```

### Providers with locale support

If your provider will need to handle different locales, it gets a bit more
complex. In the Python library [faker][] from which this package draws 
inspiration, you can create separate folders for each provider within the 
Python library. 

However, R doesn't allow this, so instead we categorize different locales for
each provider within the file names. For example, for the address provider we 
have files in the package:

- `address-provider.R`
- `address-provider-en_US.R`
- `address-provider-en_GB.R`

Where the latter two provides specific data for each locale, and the first 
file has the `AddressProvider` class that pulls in the locale specific data.

Here, we'll create a very simplified `AddressProvider` class using an 
example locale file.

```{r}
library(charlatan)
file <- system.file("examples", "address-provider-en_US.R", package = "charlatan")
source(file)
MyAddressProvider <- R6::R6Class(
  inherit = MyBaseProvider,
  'MyAddressProvider',
  lock_objects = FALSE,
  public = list(
    locale = NULL,
    city_suffixes = NULL,

    initialize = function() {
      self$locale <- 'en_us'
      self$city_suffixes <- 
        eval(parse(text = paste0("city_suffixes_", self$locale)))
    },

    city_suffix = function() {
      super$random_element(self$city_suffixes)
    }
  )
)
```

We can create an instance of the `MyAddressProvider` class by calling `$new()` on it.
It only has one method `city_suffix()`, which we can call to get a random 
city suffix.

```{r}
x <- MyAddressProvider$new()
x
x$city_suffix()
```

#### Adding a new locale

When you want to add a new locale to an existing provider, look in the `R/` folder
of the package and the locales that are available are in the file names. 

Pick one of the locale files for the provider you're extending, make a duplicate of it 
and rename the file with your new locale. Then modify the duplicate, copying the 
format but putting in place the appropriate information for the new locale.

Where the data comes from for the new locale may vary. One easy way to start may be 
porting over locales in the [faker][] Python library that are not yet in `charlatan`.

If it's a locale for which you can't easily port over from another library, you need
to get the data from a variety of sources. There are some R based packages that 
should help:

- [WikidataR][]
- [humaniformat][]
- [WikidataQueryServiceR][]


## How locale specific data are used in providers

It's a little tricky how this is done. In the `initialize()` block of each main provider 
file (e.g., `address-provider.R`) we pull in the appropriate locale specific data
based on the user input locale. For example, here's an abbreviated `initialize` block from
the `AddressProvider`:

```r
initialize = function(locale = NULL) {
  if (!is.null(locale)) {
    # check global locales
    super$check_locale(locale)
    # check address provider locales
    check_locale_(locale, address_provider_locales)
    self$locale <- locale
  } else {
    self$locale <- 'en_US'
  }

  self$city_prefixes <- parse_eval("city_prefixes_", self$locale)
}
```

A few things to note:

* if no locale is given, we by default use `en_US`
* we check that the locale given is in the allowed set
* for each data type, here just city prefixes shown, use the internal function
`parse_eval()` to pull in the data. Essentially, `parse_eval()` makes the 
string `city_prefixes_en_US`, then finds that in the package environment
and `eval()`'s it to bring the data into the R6 object in the `city_prefixes`
slot. We repeat this for each data type. The result is the user initialized 
class with locale specific data.


[faker]: https://github.com/joke2k/faker
[humaniformat]: https://github.com/Ironholds/humaniformat
[WikidataQueryServiceR]: https://cran.r-project.org/package=WikidataQueryServiceR

